Revisited hidden weight bit function

### 2024-10-27 RevisedHiddenWeight.tex
modifications all Section 3

\def\anon{1} %% set to 1 for anonymous submissions, hides acknowledgements and author names
\def\full{1} %% set to 0 for springer proceedings

@@

\begin{document}
	\title{Draft: Notes on the revised Hidden weight function}

	\titlerunning{Notes on the revised Hidden weight function}
	\author{
%		\mbox{Pierrick M\'eaux}\inst{1}
	}

	\authorrunning{ P. M\'eaux}

	\institute{
	Luxembourg university, Luxembourg\\
%		\email{pierrick.meaux@uni.lu}		
	}

@@

\begin{definition}[Walsh transform and restricted Walsh transform]\label{def:walsh_transform}
	Let $f\in \Bn{n}$ be a Boolean function, its Walsh transform $\wt{f}$ at $a \in \F_2^n$ is defined as:
	\[  \wt{f} (a) := \sum_{x \in \F_2^n} (-1)^{f(x) +  a \cdot x }.\]
	%	The Walsh support is the set $\Wsupp{f}:=\{ a\in \F_2^n\, | \, \wt{f} (a) \neq 0 \}$.
	Let $f\in \Bn{n}$, $S \subset \F_2^n$, its Walsh transform restricted to $S$ at $a \in \F_2^n$ is defined as:
	\[  \wt{f,S} (a) := \sum_{x\in S} (-1)^{f(x)+ax}.\]
	For $S=\Ekn{k}{n}$ we denote $\wt{f,\Ekn{k}{n}} (a)$ by $\wtk{f}{k}(a)$.%, and for $a= 0_{n}$ we denote $\wtk{f}{k}(a)$ as $\wtk{f}{k}(0)$.
\end{definition}

@@

\begin{definition}[Weightwise degree-$d$ functions (~\cite{DAM:MeaOza24} Definition 12)]\label{def:WdegD}
	Let $n\in \N^*$ and for $k\in[0,n]$ $\phikn{k}{n}$ denotes the indicator function of $\Ekn{k}{n}$. 
	An $n$-variable Boolean function $f$, written as $f=\sum_{k=0}^n f_k \phikn{k}{n}$, is called weightwise degree-$d$ if and only if for each $k\in [0,n]$ $f_k$ coincide with a function of degree at most $d$ over $\Ekn{k}{n}$. 
		The set of weightwise degree-$d$ functions is denoted by $\WD{d}{n}$.
\end{definition}

@@

In this part we formally define the revisited HWBF function $f$ and introduce a quantity that allows us to study its weightwise nonlinearity and balancedness.

@@

We study the restricted Walsh transform  of $f$ to bound its nonlinearity.

\begin{align*}
\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+ax}\\
&=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ ax}\\
&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ (c)x},
\end{align*}
where $c$ is the $n$-length binary vector such that $c=a+e_k+\sum_{i=n/2+1}^n e_i$ where the $e_i$ denotes the canonical vector with "1" in position $i$.

@@

\begin{definition}
Let $n\in \N$ be an even number, and $a\in \F_2^n$. We denote $d_n$ the function given by: $d_n(x)=\sum_{i \text{ odd}}^n x_i x_{i+1}$, 
and we denote by $\Dnka{n}{k}{a}$:
\[\Dnka{n}{k}{a}= \sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a x}\]
\end{definition}
The limit case of $n=0$ yields that $\Dnka 0ka=1$ if $k=0$ and $\Dnka 0ka=0$ otherwise.

With these notations we obtain a recursive relation for $\Dnka{n}{k}{a}$, and therefore for $\wtk{f}{k}(a)$, denoting $b=(a_1,\cdots,a_{n-2})$ and $y=(x_1,\cdots,x_{n-2})$, we get for any even $n\geq 2$:

@@

Then it gives three cases:
\begin{itemize}
	\item if $a_{n-1}=a_n=0$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}+2\Dnka{n}{k-1}{b}-\Dnka{n}{k-2}{b}$,
	\item if $a_{n-1} \ne a_n$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}+\Dnka{n}{k-2}{b}$,
	\item if $a_{n-1}=a_n=1$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}-2\Dnka{n}{k-1}{b}-\Dnka{n}{k-2}{b}$.
\end{itemize}

In this analysis we considered the different cases depending on the values of $a_{n-1}$ and $a_n$, but the same reasoning applies with any $a_i, a_{i+1}$ where $i$ is odd. 
Thereafter, the value of $\Dnka{n}{k}{a}$ depends only on the number of pairs of $a$ being $(0,0)$, $(1,1)$ or $(0,1)$ or $(1,0)$, and the values $\Dnka{2}{k}{a}$ for $k \in [0,2]$ and $a\in \{ (0,0),(1,1),(0,1)\}$.

The first values of $\Dnka{2}{k}{a}$ are listed in Table~\ref{tab:Dnka}

@@

The restricted Walsh transform of $f$ can the be written as \begin{align*}
\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+ax}=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ (c)x}\\
&=\Dnka{n}{k}{d},
\end{align*}
where $d$ is the permutation of $c$ obtained by sending the first $n/2$ elements to the odd positions, and the $n/2$ last ones to the even positions. Calling $\pi$ this permutation we can write $d=\pi(a)+ \pi(e_k)+ (0,1,0,1,\ldots,0,1)$. 
Accordingly, if $k\le n/2$, the vector $(\pi(e_k)+ (0,1,0,1,\ldots,0,1))$ has weight $n/2+1$, with one couple of consecutive elements (the first one with an odd index) being $(1,1)$ and all the others $(0,1)$. 
Otherwise,  when $k> n/2$, the vector $(\pi(e_k)+ (0,1,0,1,\ldots,0,1))$ has weight $n/2-1$, with one couple of consecutive elements (the first one with an odd index) being $(0,0)$ and all the others $(0,1)$.

Denoting by $b$ the vector $\pi(a)+(0,1,0,1,\ldots,0,1)$, we can write the Walsh transform of $f$ in $a$ as:
\begin{equation}\label{eqn:walsh_from_D}
    \wt{f}(a)=1+\sum_{k=1}^n\Dnka{n}{k}{b+\pi(e_k)}.
\end{equation}

In particular, for $a=0_n$ we can study the balancedness of $f$:
\begin{align*}
\wt{f}(0_n)&=\sum_{k=0}^n\Dnka{n}{k}{(0,1,0,1,\ldots,0,1)+\pi(e_k)}\\
&=\sum_{k=0}^{n/2}\Dnka{n}{k}{(1,1,0,1,\ldots,0,1)}+ \sum_{k=\frac{n}{2}+1}^{n}\Dnka{n}{k}{(0,0,0,1,\ldots,0,1)}
\end{align*}
Using the recursive relation we obtain:
\begin{align*}
\Dnka{n}{k}{(1,1,0,1,\ldots,0,1)}&= \Dnka{n-2}{k}{(1,1,0,1,\ldots,0,1)} + \Dnka{n-2}{k-2}{(1,1,0,1,\ldots,0,1)} \\
&=\sum_{i=0}^{n/2-1} \binom{n/2-1}{i} \Dnka{2}{k-2i}{(1,1)}\\
&=\binom{n/2-1}{k/2} \Dnka{2}{0}{(1,1)}+ \binom{n/2-1}{(k-1)/2} \Dnka{2}{1}{(1,1)}+ \binom{n/2-1}{k/2-1} \Dnka{2}{2}{(1,1)}\\
&=\binom{n/2-1}{k/2} -2 \binom{n/2-1}{(k-1)/2} - \binom{n/2-1}{k/2-1}\\
\end{align*}
Similarly, for $\Dnka{n}{k}{(0,0,0,1,\ldots,0,1)}$ we get:
\begin{align*}
\Dnka{n}{k}{(0,0,0,1,\ldots,0,1)}
%&= \Dnka{n-2}{k}{(0,0,0,1,\ldots,0,1)} + \Dnka{n-2}{k-2}{(0,0,0,1,\ldots,0,1)} \\
&=\sum_{i=0}^{n/2-1} \binom{n/2-1}{i} \Dnka{2}{k-2i}{(0,0)}\\
&=\binom{n/2-1}{k/2} \Dnka{2}{0}{(0,0)}+ \binom{n/2-1}{(k-1)/2} \Dnka{2}{1}{(0,0)}+ \binom{n/2-1}{k/2-1} \Dnka{2}{2}{(0,0)}\\
&=\binom{n/2-1}{k/2} +2 \binom{n/2-1}{(k-1)/2} - \binom{n/2-1}{k/2-1}\\
\end{align*}

%We can study the balancedness of $f$ using the $\Dnka{n}{k}{a}$ values.
Combining both we obtain:
\begin{align*}
\wt{f}(0_n)&=\sum_{k=0}^{n/2}\Dnka{n}{k}{(1,1,0,1,\ldots,0,1)}+ \sum_{k=\frac{n}{2}+1}^{n}\Dnka{n}{k}{(0,0,0,1,\ldots,0,1)}\\
&=\sum_{k=0}^{n/2} \left(\binom{n/2-1}{k/2} -2 \binom{n/2-1}{(k-1)/2} - \binom{n/2-1}{k/2-1} \right)\\
&\qquad+ \sum_{k=\frac{n}{2}+1}^{n}   \left(  \binom{n/2-1}{k/2} +2 \binom{n/2-1}{(k-1)/2} - \binom{n/2-1}{k/2-1} \right)\\
&=\sum_{k=0}^{n}\left( \binom{n/2-1}{k/2} - \binom{n/2-1}{k/2-1}\right) 
+ 2 \left( \sum_{k=\frac{n}{2}+1}^{n} \binom{n/2-1}{(k-1)/2} - \sum_{k=0}^{n/2} \binom{n/2-1}{(k-1)/2}\right)\\
&=2 \left( \sum_{k=\frac{n}{2}+1}^{n} \binom{n/2-1}{(k-1)/2} - \sum_{k=0}^{n/2} \binom{n/2-1}{(k-1)/2}\right)\\
&=2 \left( \sum_{i=\lceil n/4 \rceil}^{n/2-1} \binom {n/2-1}{i} - \sum_{i=0}^{\lfloor(n-2)/4 \rfloor} \binom{n/2-1}{i}\right).\\
\end{align*}

Then, depending on the parity of $n/2$ we obtain the value of $\wt{f}(0_n)$ (using that $\binom{n}{k}=\binom{n}{n-k}$):
	\[\wt{f}(0_n)= \left \{
\begin{array}{l l}
0 & \text{ if } n/2 \text{ is even, } \\
-2 \binom{n/2-1}{(n-2)/4} & \text{ if }  n/2 \text{ is odd.}
\end{array}\right.\]

@@

\subsection{Study of $\Dnka nka$ through Generating Functions}
For $a\in\F_2^n$, denote by $p=p(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})=(0,0)$, by $q=q(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})=(1,1)$, and by $r=r(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})\in\{(0,1),(1,0)\}$. We then have $p+q+r=n/2$. Consider further the integer polynomial $P_a(z)$ given by the following expression:
\[
P_a(z)=(-z^2+2z+1)^{p}\cdot(-z^2-2z+1)^{q}\cdot(z^2+1)^{r}.
\]

@@

This method is based on the identity (\ref{eqn:walsh_from_D}). For this, consider a vector $b\in\mathbb F_2^n$; we write $(p,q,r)=(p(b),q(b),r(b))$. Also, for $1\leq k\leq n$, let $b_k=b+\pi(e_k)$ and $(p_k,q_k,r_k)=(p(b_k),q(b_k),r(b_k))$. Observe that for every $1\leq k\leq n$, we have $(p_k,q_k,r_k)\in\{(p\pm1,q,r\mp 1),(p,q\pm 1,r\mp 1)\}$. Furthermore, observe that if $k\leq n/2$ and $(p_k,q_k,r_k)=(p+\alpha,q+\beta,r+\gamma)$ for $\{\alpha,\beta,\gamma\}=\{0,\pm 1\}$ with $\gamma\neq 0$, then we have $(p_{k+n/2},q_{k+n/2},r_{k+n/2})=(p+\alpha',q+\beta',r+\gamma')$ for some $\{\alpha',\beta',\gamma'\}=\{0,\pm 1\}$ with $\gamma'\neq 0$ which only depend on $\alpha$, $\beta$ and $\gamma$. Explicitly, we have $(\alpha',\beta',\gamma')=(\alpha,\beta,\gamma)$ if $\gamma=1$ and $(\alpha',\beta',\gamma')=(\beta,\alpha,\gamma)$ if $\gamma=-1$.

@@

\subsection{Algebraic immunity of $f$}

>>>>

\def\anon{0} %% set to 1 for anonymous submissions, hides acknowledgements and author names
\def\full{1} %% set to 0 for springer proceedings

@@

\begin{document}
	\title{Revisited hidden weight bit function}

	\titlerunning{Revisited Hidden weight bit function}
\author{
	Pierrick M\'eaux\inst{1},
	Tim Seuré\inst{1}, 
	Deng Tang\inst{2}
%	Pierrick M\'eaux\orcidID{0000-0001-5733-4341}\inst{1},
%	Tim Seuré\inst{1}, 
%	Deng Tang\orcidID{0000-0002-8373-9200}\inst{2}
}
\authorrunning{P. M\'eaux, T. Seuré, D. Tang}

\institute{
	University of Luxembourg, Luxembourg\\ \email{pierrick.meaux@uni.lu, tim.seure@uni.lu} 
	%\and
%		University of Luxembourg, Luxembourg\\ \email{pierrick.meaux@uni.lu}
	\and
	Shanghai Jiao Tong University, Shanghai, China.\\
	\email{dengtang@sjtu.edu.cn}
}

@@

\begin{definition}[Walsh transform and restricted Walsh transform]\label{def:walsh_transform}
	Let $f\in \Bn{n}$ be a Boolean function, its Walsh transform $\wt{f}$ at $a \in \F_2^n$ is defined as:
	\[  \wt{f} (a) := \sum_{x \in \F_2^n} (-1)^{f(x) +  a \cdot x }.\]
	%	The Walsh support is the set $\Wsupp{f}:=\{ a\in \F_2^n\, | \, \wt{f} (a) \neq 0 \}$.
	Let $f\in \Bn{n}$, $S \subset \F_2^n$, its Walsh transform restricted to $S$ at $a \in \F_2^n$ is defined as:
	\[  \wt{f,S} (a) := \sum_{x\in S} (-1)^{f(x)+a \cdot x}.\]
	For $S=\Ekn{k}{n}$ we denote $\wt{f,\Ekn{k}{n}} (a)$ by $\wtk{f}{k}(a)$.%, and for $a= 0_{n}$ we denote $\wtk{f}{k}(a)$ as $\wtk{f}{k}(0)$.
\end{definition}

@@

\begin{definition}[Weightwise degree-$d$ functions (~\cite{DAM:MeaOza24} Definition 12)]\label{def:wwdegd}
	Let $n\in \N^*$ and for $k\in[0,n]$ $\phikn{k}{n}$ denotes the indicator function of $\Ekn{k}{n}$. 
	An $n$-variable Boolean function $f$, written as $f=\sum_{k=0}^n f_k \phikn{k}{n}$, is called weightwise degree-$d$ if and only if for each $k\in [0,n]$ $f_k$ coincide with a function of degree at most $d$ over $\Ekn{k}{n}$. 
		The set of weightwise degree-$d$ functions is denoted by $\WD{d}{n}$.
\end{definition}

@@

In this part we formally define the revisited HWBF function $f$ and introduce a quantity that allows us to study its balancedness and nonlinearity.

@@

Since $f$ is the sum of a quadratic function and a weightweise affine function, it is a weightwise quadratic function. 
Using the formalism of Definition~\ref{def:wwdegd} $f$ is the WWQ function defined by:% $f_0=0$ and for $k\in [1,n]$ $f_k=x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+n/2}$.
\[f_0=0, \quad \text{ and } \quad \forall k \in [1,n], \, f_k= \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}.\]

This form allows to derive the restricted Walsh transform of $f$, which can be a useful tool to study the balancedness and bound the nonlinearity of a function.
%Maybe add examples, such as OM24, M24

%We study the restricted Walsh transform  of $f$ to bound its nonlinearity.
\begin{proposition}\label{prop:restrWT}
Let $n \in \N^*$ even, and $f$ be the function from Definition~\ref{def:revHWBF}, for all $k\in [0,n]$ its restricted Walsh transform in $a\in \F_2^n$ is given by:
\[\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},\]
where $c$ is the $n$-length binary vector such that $c=a+e_k+\sum_{i=n/2+1}^n e_i$ where the $e_i$ denotes the canonical vector with $1$ in position $i$.
\end{proposition}
\begin{proof}
%\begin{align*}
%	\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}
%	=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ a \cdot x}\\
%	&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},
%\end{align*}
%\begin{align*}
\[	\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}
	=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ a \cdot x}\\
	=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},\]
where $c$ denotes the $n$-length binary vector such that $c=a+e_k+\sum_{i=n/2+1}^n e_i$ and $e_i$ denotes the $i$-th canonical vector.
\end{proof}

%\begin{align*}
%\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+ax}\\
%&=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ ax}\\
%&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ (c)x},
%\end{align*}
%where $c$ is the $n$-length binary vector such that $c=a+e_k+\sum_{i=n/2+1}^n e_i$ where the $e_i$ denotes the canonical vector with "1" in position $i$.

@@

\begin{definition}
Let $n\in \N$ be an even number, and $a\in \F_2^n$. We denote $d_n$ the function given by: $d_n(x)=\sum_{i \text{ odd}}^n x_i x_{i+1}$, 
and we denote by $\Dnka{n}{k}{a}$ the quantity:
\[\Dnka{n}{k}{a}= \sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a x}\]
\end{definition}

We can give a recursive relation for $\Dnka{n}{k}{a}$.

\begin{proposition}\label{prop:recursiveDnka}
Let $n\in \N$ be an even number, $k$ an integer in the range $[0,n]$ and $a\in \F_2^n$. the Following holds on $\Dnka{n}{k}{a}$:
\begin{itemize}
	\item For $n=0$, $\Dnka{0}{k}{a}=1$ if $k=0$ and $\Dnka 0ka=0$ otherwise.
	\item if $n\ge 2$, denoting  $b=(a_1,\cdots,a_{n-2})$ and $y=(x_1,\cdots,x_{n-2})$, we get:
	\[\Dnka{n}{k}{a}=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b} + (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.\]
\end{itemize}

\begin{proof}
For $n\ge 2$ we can rewrite $\Dnka{n}{k}{a}$:

@@

The limit cases $\Dnka 0ka=1$ if $k=0$ and $\Dnka 0ka=0$ otherwise are obtained from the values of $\Dnka{2}{k}{a}$ given in Table~\ref{tab:Dnka}.
\end{proof}

%The limit case of $n=0$ yields that $\Dnka 0ka=1$ if $k=0$ and $\Dnka 0ka=0$ otherwise.

%With these notations we obtain a recursive relation for $\Dnka{n}{k}{a}$, and therefore for $\wtk{f}{k}(a)$, denoting $b=(a_1,\cdots,a_{n-2})$ and $y=(x_1,\cdots,x_{n-2})$, we get for any even $n\geq 2$:

We remark that it gives three different cases depending on the values of the two last elements of $a$:%Then it gives three cases:
\begin{itemize}
	\item if $a_{n-1}=a_n=0$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}+2\Dnka{n}{k-1}{b}-\Dnka{n}{k-2}{b}$,
	\item if $a_{n-1} \ne a_n$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}+\Dnka{n}{k-2}{b}$,
	\item if $a_{n-1}=a_n=1$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}-2\Dnka{n}{k-1}{b}-\Dnka{n}{k-2}{b}$.
\end{itemize}

%In this analysis we considered the different cases depending on the values of $a_{n-1}$ and $a_n$, but the same reasoning applies with any $a_i, a_{i+1}$ where $i$ is odd.

The recursive formula of Proposition~\ref{prop:recursiveDnka} gives different cases depending on the values of $a_{n-1}$ and $a_n$, however the same reasoning applies with any $a_i, a_{i+1}$ where $i$ is odd. 
Thereafter, the value of $\Dnka{n}{k}{a}$ depends only on the number of pairs of $a$ being $(0,0)$, $(1,1)$ or $(0,1)$ or $(1,0)$, and the values $\Dnka{2}{k}{a}$ for $k \in [0,2]$ and $a \in \{ (0,0),(1,1),(0,1)\}$. 
We give the first values of $\Dnka{2}{k}{a}$ in Table~\ref{tab:Dnka},  these values and Proposition~\ref{prop:recursiveDnka} are sufficient to determine any $\Dnka{n}{k}{a}$. 
In the following, we show how the Walsh transform of $f$ can be written in terms of $\Dnka{n}{k}{a}$. Then, in Section~\ref{sec:balancedness} we use Proposition~\ref{prop:recursiveDnka} to determine the balancedness of $f$, and in Section~\ref{sec:genfunc} we study the value of $\Dnka{n}{k}{a}$ using trough generating functions.

%The first values of $\Dnka{2}{k}{a}$ are listed in Table~\ref{tab:Dnka}

@@

Then, we can express the restricted Walsh transform of $f$ in terms of $\Dnka{n}{k}{a}$:

\begin{proposition}[Walsh transform of $f$]\label{prop:WT}
	Let $n \in \N^*$ even, and $f$ be the function from Definition~\ref{def:revHWBF}, for all $a\in \F_2^n$ its Walsh transform is given by:
	\[ \wt{f}(a)=1+\sum_{k=1}^n \Dnka{n}{k}{b+\pi(e_k)},\]
where $\pi$ is the permutation sending the first $n/2$ elements to the odd positions, and the $n/2$ last ones to the even positions, and $b=\pi(a)+(0,1,0,1,\ldots,0,1)$.
\end{proposition}

\begin{proof}
Using Proposition~\ref{prop:restrWT}, the restricted Walsh transform of $f$ can the be written as \[
\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+ax}=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x}=\Dnka{n}{k}{d},\]
where $d$ is the permutation of $c$ obtained by sending the first $n/2$ elements to the odd positions, and the $n/2$ last ones to the even positions.

Calling $\pi$ this permutation, since $c=a+e_k+\sum_{i=n/2+1}^n e_i$, we can write $d=\pi(c)=\pi(a)+ \pi(e_k)+ (0,1,0,1,\ldots,0,1)$. 
Accordingly, if $k\le n/2$, the vector $(\pi(e_k)+ (0,1,0,1,\ldots,0,1))$ has weight $n/2+1$, with one couple of consecutive elements (the first one with an odd index) being $(1,1)$ and all the others $(0,1)$. 
Otherwise,  when $k> n/2$, the vector $(\pi(e_k)+ (0,1,0,1,\ldots,0,1))$ has weight $n/2-1$, with one couple of consecutive elements (the first one with an odd index) being $(0,0)$ and all the others $(0,1)$.

Denoting by $b$ the vector $\pi(a)+(0,1,0,1,\ldots,0,1)$, we can write the Walsh transform of $f$ in $a$ as:
\[ \wt{f}(a)=1+\sum_{k=1}^n\Dnka{n}{k}{b+\pi(e_k)}.\]
\end{proof}

\subsection{Balancedness of $f$}\label{sec:balancedness}

In this part we determine for which values of $n$ the function $f$ is balanced, using the expression of its Walsh transform in terms of $\Dnka{n}{k}{a}$ and the recursive relation of these quantities.

\begin{theorem}[Balancedness of $f$]
	Let $n \in \N^*$ even, and $f$ be the $n$-variable function from Definition~\ref{def:revHWBF}, the following holds on its Walsh transform in $0_n$:
	\[\wt{f}(0_n)= \left \{
	\begin{array}{l l}
	0 & \text{ if } n/2 \text{ is even, } \\
	-2 \binom{n/2-1}{(n-2)/4} & \text{ if }  n/2 \text{ is odd.}
	\end{array}\right.\]

@@

\end{theorem}

\begin{proof}
We use the expression of the Walsh transform of $f$ from Proposition~\ref{prop:WT}:
%In particular, for $a=0_n$ we can study the balancedness of $f$:
%\begin{align*}
%\wt{f}(0_n)&=\sum_{k=0}^n\Dnka{n}{k}{(0,1,0,1,\ldots,0,1)+\pi(e_k)}\\
%&=\sum_{k=0}^{n/2}\Dnka{n}{k}{(1,1,0,1,\ldots,0,1)}+ \sum_{k=\frac{n}{2}+1}^{n}\Dnka{n}{k}{(0,0,0,1,\ldots,0,1)}
%\end{align*}
\[\wt{f}(0_n)=\sum_{k=0}^n\Dnka{n}{k}{(0,1,0,1,\ldots,0,1)+\pi(e_k)}=\sum_{k=0}^{n/2}\Dnka{n}{k}{(1,1,0,1,\ldots,0,1)}+ \sum_{k=\frac{n}{2}+1}^{n}\Dnka{n}{k}{(0,0,0,1,\ldots,0,1)}\]
Using the recursive relation from Proposition~\ref{prop:recursiveDnka} we obtain:
\begin{align*}
\Dnka{n}{k}{(1,1,0,1,\ldots,0,1)}&= \Dnka{n-2}{k}{(1,1,0,1,\ldots,0,1)} + \Dnka{n-2}{k-2}{(1,1,0,1,\ldots,0,1)} \\
&=\sum_{i=0}^{n/2-1} \binom{n/2-1}{i} \Dnka{2}{k-2i}{(1,1)}\\
&=\binom{n/2-1}{k/2} \Dnka{2}{0}{(1,1)}+ \binom{n/2-1}{(k-1)/2} \Dnka{2}{1}{(1,1)}+ \binom{n/2-1}{k/2-1} \Dnka{2}{2}{(1,1)}\\
&=\binom{n/2-1}{k/2} -2 \binom{n/2-1}{(k-1)/2} - \binom{n/2-1}{k/2-1}\\
\end{align*}
Similarly, for $\Dnka{n}{k}{(0,0,0,1,\ldots,0,1)}$ we get:
\begin{align*}
\Dnka{n}{k}{(0,0,0,1,\ldots,0,1)}
%&= \Dnka{n-2}{k}{(0,0,0,1,\ldots,0,1)} + \Dnka{n-2}{k-2}{(0,0,0,1,\ldots,0,1)} \\
&=\sum_{i=0}^{n/2-1} \binom{n/2-1}{i} \Dnka{2}{k-2i}{(0,0)}\\
&=\binom{n/2-1}{k/2} \Dnka{2}{0}{(0,0)}+ \binom{n/2-1}{(k-1)/2} \Dnka{2}{1}{(0,0)}+ \binom{n/2-1}{k/2-1} \Dnka{2}{2}{(0,0)}\\
&=\binom{n/2-1}{k/2} +2 \binom{n/2-1}{(k-1)/2} - \binom{n/2-1}{k/2-1}\\
\end{align*}

%We can study the balancedness of $f$ using the $\Dnka{n}{k}{a}$ values.
Combining both we obtain:
%\begin{align*}
%\wt{f}(0_n)&=\sum_{k=0}^{n/2}\Dnka{n}{k}{(1,1,0,1,\ldots,0,1)}+ \sum_{k=\frac{n}{2}+1}^{n}\Dnka{n}{k}{(0,0,0,1,\ldots,0,1)}\\
%&=\sum_{k=0}^{n/2} \left(\binom{n/2-1}{k/2} -2 \binom{n/2-1}{(k-1)/2} - \binom{n/2-1}{k/2-1} \right)\\
%&\qquad+ \sum_{k=\frac{n}{2}+1}^{n}   \left(  \binom{n/2-1}{k/2} +2 \binom{n/2-1}{(k-1)/2} - \binom{n/2-1}{k/2-1} \right)\\
%&=\sum_{k=0}^{n}\left( \binom{n/2-1}{k/2} - \binom{n/2-1}{k/2-1}\right) 
%+ 2 \left( \sum_{k=\frac{n}{2}+1}^{n} \binom{n/2-1}{(k-1)/2} - \sum_{k=0}^{n/2} \binom{n/2-1}{(k-1)/2}\right)\\
%&=2 \left( \sum_{k=\frac{n}{2}+1}^{n} \binom{n/2-1}{(k-1)/2} - \sum_{k=0}^{n/2} \binom{n/2-1}{(k-1)/2}\right)\\
%&=2 \left( \sum_{i=\lceil n/4 \rceil}^{n/2-1} \binom {n/2-1}{i} - \sum_{i=0}^{\lfloor(n-2)/4 \rfloor} \binom{n/2-1}{i}\right).\\
%\end{align*}
\begin{align*}
\wt{f}(0_n)&=\sum_{k=0}^{n/2}\Dnka{n}{k}{(1,1,0,1,\ldots,0,1)}+ \sum_{k=\frac{n}{2}+1}^{n}\Dnka{n}{k}{(0,0,0,1,\ldots,0,1)}\\
&=\sum_{k=0}^{\frac{n}{2}} \left(\binom{\frac{n}{2}-1}{\frac{k}{2}} -2 \binom{\frac{n}{2}-1}{\frac{k-1}{2}} - \binom{\frac{n}{2}-1}{\frac{k}{2}-1} \right)+ \sum_{k=\frac{n}{2}+1}^{n}   \left(  \binom{\frac{n}{2}-1}{\frac{k}{2}} +2 \binom{\frac{n}{2}-1}{\frac{k-1}{2}} - \binom{\frac{n}{2}-1}{\frac{k}{2}-1} \right)\\
&=\sum_{k=0}^{n}\left( \binom{\frac{n}{2}-1}{\frac{k}{2}} - \binom{\frac{n}{2}-1}{\frac{k}{2}-1}\right) 
+ 2 \left( \sum_{k=\frac{n}{2}+1}^{n} \binom{\frac{n}{2}-1}{\frac{k-1}{2}} - \sum_{k=0}^{\frac{n}{2}} \binom{\frac{n}{2}-1}{\frac{k-1}{2}}\right)\\
&=2 \left( \sum_{k=\frac{n}{2}+1}^{n} \binom{\frac{n}{2}-1}{\frac{k-1}{2}} - \sum_{k=0}^{n/2} \binom{\frac{n}{2}-1}{\frac{k-1}{2}}\right)
=2 \left( \sum_{i=\lceil \frac{n}{4} \rceil}^{\frac{n}{2}-1} \binom {\frac{n}{2}-1}{i} - \sum_{i=0}^{\lfloor\frac{n-2}{4}\rfloor} \binom{\frac{n}{2}-1}{i}\right).\\
\end{align*}
Then, depending on the parity of $\frac{n}{2}$ we obtain the value of $\wt{f}(0_n)$ (using that $\binom{n}{k}=\binom{n}{n-k}$):
	\[\wt{f}(0_n)= \left \{
\begin{array}{l l}
0 & \text{ if } \frac{n}{2} \text{ is even, } \\
-2 \binom{\frac{n}{2}-1}{\frac{n-2}{4}} & \text{ if }  \frac{n}{2} \text{ is odd.}
\end{array}\right.\]

Accordingly, $f$ is balanced when $n \equiv 0\mod 4$ and unbalanced when $n\equiv 2 \mod 4$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Extensive study of $\Dnka nka$ and bounds on the Walsh spectrum of $f$}

\subsection{Study of $\Dnka nka$ through Generating Functions}\label{sec:genfunc}
For $a\in\F_2^n$, denote by $p=p(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})=(0,0)$, by $q=q(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})=(1,1)$, and by $r=r(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})\in\{(0,1),(1,0)\}$. We then have $p+q+r=n/2$. Consider further the integer polynomial $P_a(z)$ given by the following expression:
\[
P_a(z)=(-z^2+2z+1)^{p}\cdot(-z^2-2z+1)^{q}\cdot(z^2+1)^{r}.
\]

@@

@@

\section{Experiments and comparisons}
